\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-11-04 k-planar.tex
updated refs

\begin{table}[h]
  \centering{
    \begin{tabular}{lccl}
    	\hline
      {Graph class} & $\tw(H)$ & $\ell$ & {Reference} \\ \hline
      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
      planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
      planar & $6^*$ & $1$ & \cite{UWY22} \\
      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
      genus $g$ & $2g+6$ & 1 & \cite{UWY22} \\
      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
      $k$-nearest-neighbour & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline
      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
    \end{tabular}
  }
  \caption{Product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
  \label{summary_table}
\end{table}

@@

\begin{thm}[\citep{DJMMUW20,UWY22,DHHW}]\label{GenusProduct}
  Every graph of Euler genus $g$ is contained in:
  \begin{compactenum}[(a)]
  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some planar graph $H$ of treewidth at most $3$ and for some path $P$.
  \end{compactenum}
\end{thm}

@@

\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite{DJMMUW20}. The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{KV12}. Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]

>>>>

\begin{table}[h]
  \centering{
    \begin{tabular}{lccl}
    	\hline
      {Graph class} & $\tw(H)$ & $\ell$ & {Reference} \\ \hline
      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
      planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
      planar & $6^*$ & $1$ & \cite{UWY22} \\
      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW22} \\
      genus $g$ & $2g+6$ & 1 & \cite{UWY22} \\
      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
      $k$-nearest-neighbour & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline
      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
    \end{tabular}
  }
  \caption{Product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
  \label{summary_table}
\end{table}

@@

\begin{thm}[\citep{DJMMUW20,UWY22,DHHW22}]\label{GenusProduct}
  Every graph of Euler genus $g$ is contained in:
  \begin{compactenum}[(a)]
  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some planar graph $H$ of treewidth at most $3$ and for some path $P$.
  \end{compactenum}
\end{thm}

@@

\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite{DJMMUW20}. The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.
To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}. Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that for each $v\in V(G)$,  \[\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.\]
